\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-03-10 forwards-and-backwards.tex
up to 6.26

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order $1/\theta$.
An instructive example is the case where $\gamma \equiv 1$ and $F \equiv 0$, so that $\mu_\theta = r$,
and the population has all individuals at a single location (so, $\eta = \delta_x$),
in which case the mean rate of change of the total population size
is $\int (r(y, N \smooth{r}(y) ) - r(x, N \smooth{r}(x))) q_\theta(x, dy)$;
the first condition of Lemma~\ref{conditions on r} would ensure this is of order $1/\theta$.

@@

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \theta\frac{u}{N} \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) 
    +
    \theta\mu_\theta(x,\eta) u .
\end{equation}
We shall write 
$$
    b_\theta(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    -
    \mu_\theta(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
and
$$
    c_\theta(x, \eta)
    :=
    \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) ,
$$
which captures the local rate of production of successful offspring.
Recall from equation~\eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu_\theta(x,\eta))$,
and so 
\begin{equation*} 
b_\theta(x, \eta)
=    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) 
	q_\theta(x, dy)
    + F(x, \eta). 
\end{equation*}
Under Assumptions~\ref{def:model_setup}, as $\theta \to \infty$
$c_\theta(x, \eta)$ will tend to $\alpha \gamma(x, \eta) r(x, \eta)$
and $b_\theta(x, \eta)$ will tend to
\begin{equation}
\label{eqn:b_limit}
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) .
\end{equation}

@@

We can now define the spatial-level process
explicitly as a solution to a martingale problem,
whose generator is just the sum of
\eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.
We need some notation. 
Write ${\mathcal C}={\mathcal C}(\IR^d\times [0,\infty)$ for the point measures 
on $\IR^d\times [0,\infty)$ and ${\mathcal C}_N$ for the subset consisting of 
counting measures on $\IR^d\times [0,N]$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$, define the generator $A^N$
by
\begin{equation}
	\label{eqn:lookdown_generator}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\IR^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\IR^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation}
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ is as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$
as before. 
Given $\lp_0 \in {\mathcal C}_N$, 
we say that a ${\mathcal D}([0,\infty), {\mathcal C}_N)$-valued process 
$(\xi_t)_{t \ge 0}$
   is a solution to the $(A^N, \lp_0)$ martingale problem
    if 
$f(\lp_t) - f(\lp_0)-\int_0^t A^N f(\lp_s) ds$ is a martingale (with 
respect to the natural filtration) for all 
test functions $f$ 
    as defined above.
\end{definition}

@@

Elsewhere we have written $\eta^N$ for the spatial projection of the lookdown process $\lp^N$,
although $\eta^N$ was defined first as a solution to a martingale problem.
For the purposes of this result --
the application of the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2018} --
we must disambiguate these, and so define the spatial projection
maps $\gamma^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\gamma^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\gamma: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\gamma(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a sigma-field $\mathcal{F}$,
we say that \emph{$\lp$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle F_f^N, \gamma(\lp) \rangle} ,
\end{align}
where
$$ F_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
(In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of $\gamma(\lp)$.)
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \gamma(\lp)(dx) \rangle} .
\end{align}
(In other words, $\lp$ is conditionally Poisson with Cox measure $\gamma(\lp) \times \Lambda$.)

\begin{proposition}
    \label{thm:mmt_application}
    Let $\mu_0$ be a probability distribution on $\measures$,
    and $\nu_0^N$ a probability distribution on $\mathcal{M}(\IR^d \times [0, N])$
    such that a draw $\lp^N_0$ from $\nu_0^N$ is conditionally uniform given $\gamma(\lp^N_0)$
    in the sense of \eqref{eqn:conditionally_uniform}.
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\mu_0$
    then there exists a solution $\lp^N$
    of the martingale problem of Definition~\ref{defn:lookdown_mgale}
    with initial distribution $\nu_0^N$
    such that $\eta^N = \gamma \cdot \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, $\lp^N$ is conditionally uniform given given $\mathcal{F}^{\eta}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

    \comment{Theorem~\ref{thm:nonlocal_convergence}
        isn't written as ``the $(A, \mu)$ martingale problem is'';
        rephrase here or there?}

    Similarly,
    let $\mu_0$ be a probability distribution on $\measures$,
    and $\nu_0$ a probability distribution on $\mathcal{M}(\IR^d \times [0, \infty))$
    such that a draw $\lp_0$ from $\nu_0$ is conditionally Poisson given $\gamma(\lp_0)$
    in the sense of \eqref{eqn:conditionally_poisson}.
    If $\tilde \eta$ is a solution
    of the limiting martingale problem of Theorem~\ref{thm:nonlocal_convergence}
    with initial distribution $\mu_0$
    then there exists a solution $\lp$
    of the martingale problem of of Definition~\ref{defn:limiting_lookdown_mgale}
    with initial distribution $\nu_0$
    such that $\eta = \gamma \cdot \lp$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta$.
    Furthermore, $\lp_t$ is conditionally Poisson given $\mathcal{F}^{\eta}_t$
    in the sense of \eqref{eqn:conditionally_poisson}.
\end{proposition}

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This is the content of the Markov Mapping Theorem applied to our situation,
    \citet[Thm A.2 in][]{etheridge/kurtz:2018}.
    In the notation of that theorem, $\gamma$ is as above,
    $\alpha$ is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case),
    and we need a continuous $\psi^N(\lp)$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f \in \mathcal{D}(A^N)$ (and similarly a function $\psi$ for $A$).
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    which we precisely state, and verify, in Appendix~\ref{sec: Markov Mapping Theorem Application}.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C + u |F(x, \eta)|, \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the mostly troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align} \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align}
    The second line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the third line provides us with the required bound,
    absorbing $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and $\langle |F(x, \eta) u e^{-u}|, \lp \rangle$ into $\psi(\lp)$.
    Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    for a given $\lp$ (and hence $\eta = \gamma(\lp)$),
    $\sup_x |F(x, \eta)|$ is finite.
    Therefore, we may bound \eqref{eqn:part_of_psi} by
    $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    To do this we must restrict $\lp$ to lie in the space of measures
    for which this is finite;
    however, this is a completely metrizable space
    \citep[see Condition 2.1 in][]{etheridge/kurtz:2018},
    and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

The lookdown processes have been carefully constructed so that
past observations about the spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels. In other words, the dynamics of the lookdown
process preserve the conditionally uniform (resp.~conditionally Poisson)
structure.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
in Appendix~\ref{sec: Markov Mapping Theorem Application},
\comment{Refer more precisely}
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2019}
tells us that by ``removing labels'' in this way we recover the Markov
process describing our population.
In particular, for any %population process $(\eta^N_t)_{t \ge 0}$
%(i.e.,
solution to the martingale problem of Definition~\ref{defn:mgale_construction})
% of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
%(i.e., 
solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
are independently and uniformly distributed on $[0, N]$.
Similarly, in the scaling limit:
for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
Theorem~\ref{nonlocal_convergence}
%there is a lookdown process $(\lp^N_t)_{t \ge 0}$
there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
$\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

Now that the lookdown process has been defined,
we can present the main convergence theorem that is analogous
to Theorem~\ref{thm:nonlocal_convergence} for the population process.

\comment{the next theorem needs rewriting along the lines of Thm A.12 in 
Kurtz and Rodrigues.}

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ be as defined in Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Suppose also that $\eta^N_0 \to \eta_0$ (as measures on $\IR^d$),
    and that for each $N$, $\lp^N_0(dx, du)$ is conditionally Poisson on $\IR^d \times [0, N]$
    with Cox measure $\eta^N_0(dx) \times du$.
    Then, the sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson with Cox measure $\eta_0(dx) \times du$,
    that is a solution to the martingale problem~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

@@

The main interest in using a lookdown construction for our population
processes is that it allows us to retain information about the 
relatedness of individuals as we pass to the infinite population limit.
In order to exploit this, in this section we write down stochastic 
equations for the locations and levels of individuals in the lookdown
representation of the prelimiting model.
We will then be able to pass to the scaling limit.
This provides an explicit description of the solution to
the limiting martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}
which will enable us to identify all 
individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of ``tracers'', popular in statistical physics and used in population
genetics by a number of authors including 
\cite{hallatschek/nelson:2008, durrett/fan:2016, biswas/etheridge/klimek:2018}.

We will construct the process using a Ulam-Harris indexing scheme.
First, we assign each extant individual a unique label from $\IN_{\ge 0}$.
Suppose an individual with label $a$ and level $u$ reproduces,
and as a result there are two individuals, one with level $u$ and one with a new level $u_1 > u$.
The parent individual, previously labeled $a$, might be assigned either level.
We will track chains of descendant individuals forwards through time
following levels, rather than individuals, and will call this a \emph{line of descent}.
So, after reproduction, we give a new label to \emph{only} the individual
that is given the new level $u_1$,
retaining the label $a$ for the individual with the old level $u$.
In this way, at each birth event, a unique label is 
assigned to the resulting individual with the higher level,
and the label of an individual may change throughout its lifetime.

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order $1/\theta$.
An instructive example is the case where $\gamma \equiv 1$ and $F \equiv 0$, so that $\mu_\theta = r$,
and the population has all individuals at a single location (so, $\eta = \delta_x$),
in which case the mean rate of change of the total population size
is $\int (r(y, N \smooth{r}(y) ) - r(x, N \smooth{r}(x))) q_\theta(x, dy)$;
the first condition of Lemma~\ref{lem:conditions on r} would ensure this is of order $1/\theta$.

@@

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \theta\frac{u}{N} \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) 
    +
    \theta\mu_\theta(x,\eta) u .
\end{equation}
We shall write 
\begin{align} \label{b_defn}
    b_\theta(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    -
    \mu_\theta(x,\eta)
    \right) ,
\end{align}
which captures the local net difference between reproduction and death,
and
\begin{align} \label{c_defn}
    c_\theta(x, \eta)
    :=
    \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) ,
\end{align}
which captures the local rate of production of successful offspring.
Recall from equation~\eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu_\theta(x,\eta))$,
and so 
\begin{equation*} 
b_\theta(x, \eta)
=    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) 
	q_\theta(x, dy)
    + F(x, \eta). 
\end{equation*}
Under Assumptions~\ref{def:model_setup}, as $\theta \to \infty$
$c_\theta(x, \eta)$ will tend to $\alpha \gamma(x, \eta) r(x, \eta)$
and $b_\theta(x, \eta)$ will tend to
\begin{equation}
\label{eqn:b_limit}
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) .
\end{equation}

@@

We can now define the spatial-level process
explicitly as a solution to a martingale problem,
whose generator is just the sum of
\eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.
We need some notation. 
Write ${\mathcal C}={\mathcal C}(\IR^d\times [0,\infty)$ for the point measures 
on $\IR^d\times [0,\infty)$ and ${\mathcal C}_N$ for the subset consisting of 
counting measures on $\IR^d\times [0,N]$.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$, define the generator $A^N$
by
\begin{equation}
	\label{eqn:lookdown_generator}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\IR^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\IR^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation}
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ is as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$
as before. 
Given $\lp_0 \in {\mathcal C}_N$, 
we say that a ${\mathcal D}([0,\infty), {\mathcal C}_N)$-valued process 
$(\xi_t)_{t \ge 0}$
   is a solution to the $(A^N, \lp_0)$ martingale problem
    if 
$f(\lp_t) - f(\lp_0)-\int_0^t A^N f(\lp_s) ds$ is a martingale (with 
respect to the natural filtration) for all 
test functions $f$ 
    as defined above.
\end{definition}

@@

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2019}
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a sigma-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle F_f^N, \kappa(\lp) \rangle} ,
\end{align}
where
$$ F_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of $\kappa(\lp)$.
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform given given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

    Similarly,
    if $\tilde \eta$ is a solution
    of the limiting martingale problem of Theorem~\ref{thm:nonlocal_convergence}
    with initial distribution $\eta_0$
    then there exists a solution $\lp$
    of the martingale problem of of Definition~\ref{defn:limiting_lookdown_mgale}
    such that $\eta = \kappa \circ \lp$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta$.
    Furthermore, $\lp_t$ is conditionally Poisson given $\mathcal{F}^{\eta}_t$
    in the sense of \eqref{eqn:conditionally_poisson}.
\end{proposition}

@@

Now we can present the main convergence theorem that is analogous
to Theorem~\ref{thm:nonlocal_convergence} for the population process.

\comment{the next theorem needs rewriting along the lines of Thm A.12 in Kurtz and Rodrigues.}
\comment{TODO: it's been rewritten some; but needs checking against that Thm yet}

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ satisfy Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Let $\eta_0^N = \kappa(\lp_0^N)$ and
    suppose also that $\eta^N_0 \to \eta_0$ (as finite measures on $\IR^d$),
    and that for each $N$, $\lp^N_0$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}.
    Then, the sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson given $\eta_0 = \kappa(\lp_0)$
    in the sense of \eqref{eqn:conditionally_poisson},
    that is a solution to the martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

@@

The main interest in using a lookdown construction for our population
processes is that it allows us to retain information about the 
relatedness of individuals as we pass to the infinite population limit.
In order to exploit this, in this section we write down stochastic 
equations for the locations and levels of individuals in the prelimiting lookdown model.
We will then be able to pass to the scaling limit.
This provides an explicit description of the solution to
the limiting martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}
which will enable us to identify all 
individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of ``tracers'', popular in statistical physics and used in population
genetics by a number of authors including 
\citet{hallatschek/nelson:2008}, \citet{durrett/fan:2016}, and \citet{biswas/etheridge/klimek:2018}.

We will construct the process using a Ulam-Harris indexing scheme.
First, we assign each individual alive at time 0 a unique label from $\IN$.
Suppose an individual with label $a$ and level $u$ reproduces,
and as a result there are two individuals, one with level $u$ and one with a new level $u_1 > u$.
The parent individual, previously labeled $a$, might be assigned either level.
We will track chains of descendant individuals forwards through time by
following levels, rather than individuals, and will call this a \emph{line of descent}.
So, after reproduction, we give a new label to \emph{only} the individual
that is given the new level $u_1$,
retaining the label $a$ for the individual with the old level $u$.
In this way, at each birth event, a unique label is 
assigned to the resulting individual with the higher level,
and the label of an individual may change throughout its lifetime.

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN^k$,
let $\Pi_a$ be an independent Poisson process on 
$[0, \infty)^2 \times \IR^d \times \{0,1\}$.
The mean measure of each $\Pi_a$ is a product of Lebesgue measure on $[0, \infty)^2$,
the density of the standard Gaussian on $\IR^d$, and 
$(\delta_0 + \delta_1)/2$ on $\{0, 1\}$.
It will also be convenient
to suppose that for each label $a$ we have an enumeration of the points in $\Pi_a$,
so we may refer to ``the $j^\text{th}$ point in $\Pi_a$'',
although the precise order of this enumeration is irrelevant.
If $(\tau, v, z, \kappa)$ is the $j^\text{th}$ point in $\Pi_a$,
then $\tau$ will determine a possible birth time,
$v$ will determine the level of the offspring,
$z$ will determine the spatial displacement of the offspring relative to the parent,
$\kappa$ will be used to determine whether parent or offspring is assigned the new level,
and the new label produced will be $a \concat j$,
i.e., the label $a$ with $j$ appended
(so, if $a = (a_1, \ldots, a_k)$ then $a \concat j = (a_1, \ldots, a_k, j)$).
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a (possibly infinite) death time $\sigma_a$, when its level first hits $N$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$, respectively.
Furthermore, define
\begin{align*}
    \eta^N_t = \frac{1}{N} \sum_{a : \tau_a \le t < \sigma_a} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp^N_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} .
\end{align*}
